---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
  # , cache = TRUE
)
devtools::load_all()
```

# vivrcity

<!-- badges: start -->
<!-- badges: end -->

The goal of vivrcity is to provide a simple R interface to the [Vivacity Labs API](https://docs.vivacitylabs.com/).

## Installation

You can install the development version of vivrcity from [GitHub](https://github.com/) with:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("Robinlovelace/vivrcity")
```

## Setup

You need a Vivacity API key to use this package.
Save it in your `.Renviron` file:

```
VIVACITY_API_KEY=your_api_key_here
```

## Usage

This example demonstrates how to retrieve metadata, counts, and visualize traffic data.

Load the package:

```{r, eval=FALSE}
library(vivrcity)
```

<details>
<summary>Development</summary>

Or for developing, clone the repo with the following command:

```sh
gh repo clone Robinlovelace/vivrcity
```

See [GitHub CLI installation instructions](https://github.com/cli/cli#installation) for how to install `gh`.


And then open the folder in RStudio/VSCode/your favourite IDE and run:

```{r}
devtools::load_all()
```

</details>


## Countline metadata

```{r example, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

# Get countline metadata (returns sf object)
metadata_sf <- get_countline_metadata()
names(metadata_sf)
```

The countline dataset is represented as linestrings.
For visualisation it can be useful to convert to points:

```{r}
metadata_points <- metadata_sf |>
  sf::st_centroid()
```

You can then visualise the data with your favourite mapping package, e.g.:

```{r, eval=FALSE}
library(tmap)
tmap_mode("view")
qtm(metadata_points)
```

Note that each device typically has 2 countlines (e.g. `S38_in` and `S38_out`):

```{r}
metadata_sf$id |>
  duplicated() |>
  summary()
# for name:
metadata_sf$name |>
  duplicated() |>
  summary()
# simplified name:
metadata_sf$sensor_name |>
  duplicated() |>
  summary()
```

## Getting counts

```{r counts}
# Sample 3 random countlines and rename to A, B, C
set.seed(2025)
sampled_metadata <- metadata_sf |>
  slice_sample(n = 3) |>
  mutate(sensor = c("A", "B", "C"))
sampled_ids <- sampled_metadata$id
id_lookup <- setNames(sampled_metadata$sensor, sampled_metadata$id)

# Get counts for a week in 2025
from_time <- as.POSIXct("2025-12-01", tz = "UTC")
to_time <- as.POSIXct("2025-12-17", tz = "UTC")

# Get counts by mode/class (default with get_counts)
counts <- get_counts(sampled_ids, from = from_time, to = to_time)

names(counts)

# Plot pedestrians vs cyclists
counts |>
  filter(class %in% c("pedestrian", "cyclist", "car")) |>
  mutate(sensor = id_lookup[id]) |>
  group_by(sensor, class, day = as.Date(from)) |>
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = day, y = count, color = class)) +
  geom_line() +
  facet_wrap(~sensor, ncol = 1) +
  labs(
    x = "Date",
    y = "Count",
    color = "Mode"
  ) +
  theme_minimal()
```

## Getting speeds

Note: this will fail if the sensors don't have speed recording enabled:

```{r, eval=FALSE}
# Get speeds (function accepts vector of IDs)
speeds_df <- get_countline_speed(sampled_ids, from = from_time, to = to_time)
```

## Aggregating data

Sometimes you might want to aggregate data from multiple countlines that belong to the same sensor (e.g. `S38_in` and `S38_out`).
The package provides functions to aggregate both counts and metadata.

```{r aggregation}
# Aggregate counts (sums across directions for same ID)
# Note: get_counts(..., aggregate = TRUE) does this automatically
aggregated_counts <- aggregate_counts(counts)
nrow(aggregated_counts) / nrow(counts)
names(aggregated_counts)
```

Aggregate metadata to one row per sensor:

```{r aggregate-metadata}
# Aggregate metadata (returns sf object with one row per sensor)
# Note: get_countline_metadata(aggregate = TRUE) does this automatically
metadata_aggregated <- aggregate_metadata(metadata_sf)
names(metadata_aggregated)
nrow(metadata_sf)
nrow(metadata_aggregated)
summary(aggregated_counts$sensor_name %in% metadata_aggregated$sensor_name)
```

Let's plot the results for a single sensor:

```{r single-sensor}
# Filter to single sensor
single_sensor_id <- aggregated_counts$sensor_name[1]
metadata_sensor_level <- metadata_aggregated |>
  filter(sensor_name == single_sensor_id) |>
  sf::st_centroid()
metadata_count_level <- metadata_sf |>
  filter(sensor_name == single_sensor_id) 
# Verify countlines belong to same road:
metadata_count_level$name
plot(metadata_count_level$geometry, main = paste("Sensor", single_sensor_id))
plot(metadata_sensor_level$geometry, col = "red", pch = 19, add = TRUE)
```

```{r summary, eval=FALSE, echo=FALSE}
# Summary statistics
summary_stats <- counts_df |>
  group_by(sensor) |>
  summarise(
    observations = n(),
    total_count = sum(count, na.rm = TRUE)
  )
summary_stats

# Plot traffic counts over time
ggplot(counts_df, aes(x = from, y = count, color = sensor)) +
  geom_line() +
  labs(
    title = "Traffic Counts (Dec 2025)",
    x = "Time",
    y = "Total Vehicles",
    color = "Sensor"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r exploratory, eval=FALSE, echo=FALSE}
# Exploring the data
highest_count <- counts_df |>
  slice_max(count)

highest_count

# Let's plot the associated sensor:
highest_count_id <- highest_count$id
highest_count_sf <- metadata_sf |>
  filter(id == highest_count_id)
library(tmap)
tmap_mode("view")
qtm(highest_count_sf)
```


## Next Steps

This package currently implements a subset of the [Vivacity Traffic Monitoring API](https://docs.vivacitylabs.com/). Future development could include:

**Additional Countline Endpoints:**
- Journey times (`/countline/journey_times`) - requires ANPR-enabled sensors

**Zone-Based Data:**
- Zonal speeds (`/zone/speeds`)
- Occupancy data (`/zone/occupancy`)  
- Turning movements (`/zone/turning_movements`)
- Dwell times (`/zone/dwell_times`)

**Other Features:**
- Viewpoint metadata and data availability
- Zone metadata
- Annotations support
- V2 API endpoints for projects and positions

**Improvements:**
- Better error handling with informative messages
- Automatic pagination for large queries
- Caching of metadata requests
- Support for all vehicle/pedestrian class types

Contributions welcome! See the [API documentation](https://docs.vivacitylabs.com/) for endpoint details.

