---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vivrcity

<!-- badges: start -->
<!-- badges: end -->

The goal of vivrcity is to provide a simple R interface to the [Vivacity Labs API](https://docs.vivacitylabs.com/).

## Installation

You can install the development version of vivrcity from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("Robinlovelace/vivrcity")
```

## Setup

You need a Vivacity API key to use this package.
Save it in your `.Renviron` file:

```
VIVACITY_API_KEY=your_api_key_here
```

## Example

This example demonstrates how to retrieve metadata, counts, and visualize traffic data.

Load the package:

```{r, eval=FALSE}
library(vivrcity)
```

Or for developing, clone the repo with

```sh
gh repo clone Robinlovelace/vivrcity
```

And then open the folder in RStudio/VSCode/your favourite IDE and run:

```{r}
devtools::load_all()
```


```{r example, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

# Get countline metadata (returns sf object)
metadata_sf <- get_countline_metadata()
names(metadata_sf)
```

```{r}
# Sample 3 random countlines and rename to A, B, C
set.seed(2025)
sampled_metadata <- metadata_sf |>
  slice_sample(n = 3) |>
  mutate(sensor = c("A", "B", "C"))
sampled_ids <- sampled_metadata$id
id_lookup <- setNames(sampled_metadata$sensor, sampled_metadata$id)

# Get counts for a week in 2024
from_time <- as.POSIXct("2024-01-01", tz = "UTC")
to_time <- as.POSIXct("2024-01-08", tz = "UTC")

counts_df <- get_countline_counts(sampled_ids, from = from_time, to = to_time) |>
  mutate(sensor = id_lookup[id])
```

Note: this will fail if the sensors don't have speed recording enabled:

```{r, eval=FALSE}
# Get speeds (function accepts vector of IDs)
speeds_df <- get_countline_speed(sampled_ids, from = from_time, to = to_time)
```

We can visualise the counts:

```{r summary}
# Summary statistics
summary_stats <- counts_df |>
  group_by(sensor) |>
  summarise(
    observations = n(),
    total_count = sum(count, na.rm = TRUE)
  )
summary_stats

# Plot traffic counts over time
ggplot(counts_df, aes(x = from, y = count, color = sensor)) +
  geom_line() +
  labs(
    title = "Traffic Counts (Jan 2024)",
    x = "Time",
    y = "Total Vehicles",
    color = "Sensor"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
